---
format: gfm
bibliography: references.bib
suppress-bibliography: true
---

```{r}
#| label: setup
#| echo: false
#| output: asis
yaml_metadata <- rmarkdown::yaml_front_matter(here::here(
  "manuscript",
  "spruce_shark_yanagimoto.qmd"
))

cat("# ", yaml_metadata$title, "\n\n")
```

[![arXiv](https://img.shields.io/badge/arXiv-2602.19798-b31b1b.svg)](https://arxiv.org/abs/2602.19798)

[Kazuharu Yanagimoto](https://kazuyanagimoto.com/)


## Abstract

```{r}
#| label: abstract
#| echo: false
#| output: asis
cat(yaml_metadata$abstract)
```

## Replication

### Simple replication (Julia Only)

1. Resotre the Julia environment with `julia --project=.` and `] instantiate`
1. Run `julia --project=. Julia/estimation.jl` to replicate the main results in the paper. The output will be saved in `output/`.
1. Run `julia --project=. Julia/benchmark.jl` to replicate the computational benchmarks. The output will be saved in `output/`.

### Full replication (Julia, R, and Quarto)

1. Resotre the Julia environment with `julia --project=.` and `] instantiate`
1. Set `PATH=to/julia/bin:$PATH` in `.Renviron` to make the Julia executable available to R
1. Restore the R environment with `renv::restore()`
1. Run `targets::tar_make()` to execute the full pipeline, including estimation, figure and table generation, and the manuscript rendering. The manuscript will be saved as `manuscript/spruce_shark_yanagimoto.pdf`.

## Why is this project named "Spruce Shark🦈"?

I like Dr. Andrew Heiss' [idea](https://github.com/andrewheiss/testy-turtle) for a project name.
So I set the seed at `260216` (the day I opened the repository), and it produces a codename "Spruce Shark".

```{r}
#| label: codename
library(codename)
codename_message()

codename(seed = 260216, type = "ubuntu")
```

## Licenses

**Text and figures:** All prose and images in the manuscript are copyright of the author. The SSRN preprint is distributed under the SSRN license terms.

**Code:** All code is licensed under the [MIT License](LICENSE.md).
